---
title: "Values from Text — Reproducible Notebooks"
---

This site renders the analysis notebooks that produce the various analyses, 
figures, tables, and in-text statistical reporting that appears in the 
manuscript. 

# General Methods

## Survey Flow

```{r, child = "_notebooks/text/0_survey_flow.qmd"}
```

## Stimuli Descriptives

```{r, child = "_notebooks/text/3_appendix_A.qmd"}
```

# Pilot Study

## Demographics

```{r, child = "_notebooks/participant/00_descriptives/01_pilot_demographics.qmd"}
```

```{r, child = "_notebooks/participant/00_descriptives/02_pilot_lyrics_questions.qmd"}
```

## Quiz results

```{r, child = "_notebooks/participant/01_quiz_results/01_quiz_results.qmd"}
```

## Subjectivity & confidence

```{r, child = "_notebooks/participant/02_subjectivity_confidence/02_pilot_subjectivity_confidence.qmd"}
```

## Rater number estimation & Inter-rater reliability

```{r, child = "_notebooks/participant/03_inter_rater_reliability/03_pilot_rater_number_estimation.qmd"}
```

# Main Study

## Descriptives

```{r, child = "_notebooks/participant/00_descriptives/03_main_descriptives.qmd"}
```

## Subjectivity & confidence

```{r, child = "_notebooks/participant/02_subjectivity_confidence/02_main_subjectivity_confidence.qmd"}
```

## Inter-rater reliability

```{r, child = "_notebooks/participant/03_inter_rater_reliability/05_wave_2_reliabilities.Rmd"}
```

## Multi-Dimensional Scaling

```{r, child = "_notebooks/participant/04_mds_plots/2_MDS_plots.Rmd"}
```

## Preliminary analysis

```{r, child = "_notebooks/participant/05_variable_importance/1_preliminary_analysis.Rmd"}
```

## Lyric model selection

```{r, child = "_notebooks/participant/05_variable_importance/2_lyric_model_selection.Rmd"}
```

## Lyric model analysis

```{r, child = "_notebooks/participant/05_variable_importance/4_lyric_model_analysis.Rmd"}
```

## Speech model selection

```{r, child = "_notebooks/participant/05_variable_importance/3_speech_model_selection.Rmd"}
```

## Speech model analysis

```{r, child = "_notebooks/participant/05_variable_importance/5_speech_model_analysis.Rmd"}
```

# Grounding

```{r, child = "_notebooks/participant/05_variable_importance/6_RRA_descriptives.Rmd"}
```
