---
title: "Values from Text — Reproducible Notebooks"
---

This site renders the analysis notebooks in the order their outputs appear in the
manuscript. Each section below is a rendered notebook. Data-preparation notebooks
are noted by path but not executed on render.

# General Methods

## Survey Flow

```{r, child = "_notebooks/text/0_survey_flow.qmd"}
```

# Appendix A: Stimuli

## Genre tag preparation

Prep only (requires live Spotify/Last.fm credentials; output cached):
`_notebooks/text/1_lyric_get_genre_tags_api.Rmd`

## Bin labels

Prep only: `_notebooks/text/2_lyric_bin_labels.Rmd`

## Stimuli screening

```{r, child = "_notebooks/text/3_appendix_A.qmd"}
```

# Participant Analyses

## Descriptives — pilot demographics

```{r, child = "_notebooks/participant/00_descriptives/01_pilot_demographics.qmd"}
```

## Descriptives — pilot lyric questions

```{r, child = "_notebooks/participant/00_descriptives/02_pilot_lyrics_questions.qmd"}
```

## Descriptives — main

```{r, child = "_notebooks/participant/00_descriptives/03_main_descriptives.qmd"}
```

## Quiz results

```{r, child = "_notebooks/participant/01_quiz_results/01_quiz_results.qmd"}
```

## Subjectivity & confidence — pilot

```{r, child = "_notebooks/participant/02_subjectivity_confidence/02_pilot_subjectivity_confidence.qmd"}
```

## Subjectivity & confidence — main

```{r, child = "_notebooks/participant/02_subjectivity_confidence/02_main_subjectivity_confidence.qmd"}
```

## Inter-rater reliability — pilot rater number estimation

```{r, child = "_notebooks/participant/03_inter_rater_reliability/03_pilot_rater_number_estimation.qmd"}
```

## Inter-rater reliability — wave 2

```{r, child = "_notebooks/participant/03_inter_rater_reliability/05_wave_2_reliabilities.Rmd"}
```

## MDS plots

```{r, child = "_notebooks/participant/04_mds_plots/2_MDS_plots.Rmd"}
```

## Variable importance — preliminary analysis

```{r, child = "_notebooks/participant/05_variable_importance/1_preliminary_analysis.Rmd"}
```

## Variable importance — lyric model selection

```{r, child = "_notebooks/participant/05_variable_importance/2_lyric_model_selection.Rmd"}
```

## Variable importance — speech model selection

```{r, child = "_notebooks/participant/05_variable_importance/3_speech_model_selection.Rmd"}
```

## Variable importance — lyric model analysis

```{r, child = "_notebooks/participant/05_variable_importance/4_lyric_model_analysis.Rmd"}
```

## Variable importance — speech model analysis

```{r, child = "_notebooks/participant/05_variable_importance/5_speech_model_analysis.Rmd"}
```

## Variable importance — RRA descriptives

```{r, child = "_notebooks/participant/05_variable_importance/6_RRA_descriptives.Rmd"}
```
