---
title: "R Notebook"
---

```{r}
library("here")         # file logistics
library("lme4")         # mixed effects metric
library("lmerTest")     # p values
library("tidyverse")    # code logistics
library("broom.mixed")  # model outputs as data frames
library("emmeans")      # estimated marginal means
library("patchwork")    # merge plots
library("report")       # apa reporting
library("gt")           # latex tables
library("effectsize")   # partial eta sq calculations

# TO "CENTER" CATEGORICAL PREDICTORS, set the coding to contr.sum (deviations from the average)
#options(scipen = 999, contrasts = c("contr.sum", "contr.poly"))
path <- here("_data", "_intermediary_data", "models")
```


```{r}
# read in data
wave_2_dfs <- readRDS(here("_data", "_intermediary_data", "wave_2_dfs.RDS"))
```

# speeches: model selection

```{r}
# # Ethnicity Mods
# speeches1a_mod <- lmer(rating ~ value + Ethnicity * party + (1|item_ID) + (1 | participant_ID),
#   data = wave_2_dfs$wave_2_speeches, REML = F)
#
# speeches2a_mod <- lmer(rating ~ value * Ethnicity * party + (1|item_ID) + (1 | participant_ID),
#   data = wave_2_dfs$wave_2_speeches, REML = F)
#
# # Spectrum Mods
# speeches1b_mod <- lmer(rating ~ value + spectrum * party + (1|item_ID) + (1 | participant_ID),
#   data = wave_2_dfs$wave_2_speeches, REML = F)
#
# speeches2b_mod <- lmer(rating ~ value * spectrum * party + (1|item_ID) + (1 | participant_ID),
#   data = wave_2_dfs$wave_2_speeches, REML = F)
#
# # Combined Mods
# speeches3_mod <- lmer(rating ~ value * Ethnicity * spectrum + party + (1|item_ID) + (1 | participant_ID),
#   data = wave_2_dfs$wave_2_speeches, REML = F)
#
# speeches4_mod <- lmer(rating ~ value * Ethnicity * spectrum * party + (1|item_ID) + (1 | participant_ID),
#   data = wave_2_dfs$wave_2_speeches, REML = F)
#
# speeches_a_mods <- list(
#   speeches1a_mod = speeches1a_mod,
#   speeches2a_mod = speeches2a_mod,
#   speeches3_mod  = speeches3_mod,
#   speeches4_mod  = speeches4_mod)
#
# speeches_b_mods <- list(
#   speeches1b_mod = speeches1b_mod,
#   speeches2b_mod = speeches2b_mod,
#   speeches3_mod  = speeches3_mod,
#   speeches4_mod  = speeches4_mod)
#
# saveRDS(speeches_a_mods, here(path, "speeches_a_mods.RDS"))
# saveRDS(speeches_b_mods, here(path, "speeches_b_mods.RDS"))
```

```{r}
speeches_a_mods <- readRDS(here(path, "speeches_a_mods.RDS"))
speeches_b_mods <- readRDS(here(path, "speeches_b_mods.RDS"))

speeches1a_mod <- speeches_a_mods[["speeches1a_mod"]]
speeches1b_mod <- speeches_b_mods[["speeches1b_mod"]]
```

## helpers

```{r}
# take sequential series of models,
# extract fit indices,
# compare each model to previous in series

model_comparison_table <- function(model_list) {

  # Ensure named list
  if (is.null(names(model_list))) {
    names(model_list) <- paste0("mod", seq_along(model_list))
  }

  n_models <- length(model_list)

  # 1. Pseudo R2 (Nakagawa)
  r2_df <- purrr::map_df(model_list, ~ as.data.frame(performance::r2_nakagawa(.x)))

  # 2. AIC
  aic_vals <- sapply(model_list, AIC)

  # 3. Formulas
  formulas <- purrr::map_chr(model_list, ~ paste(deparse(formula(.x)), collapse = " "))

  # 4. Sequential LRT
  lrt_chisq <- numeric(n_models)
  lrt_df    <- numeric(n_models)
  raw_p     <- numeric(n_models)

  # first is always NA, as there is nothing to compare it to
  lrt_chisq[1] <- NA
  lrt_df[1]    <- NA
  raw_p[1]     <- NA

  for (i in 2:n_models) {
    # anova for previous model and current model
    a <- suppressWarnings(anova(model_list[[i-1]], model_list[[i]]))
    lrt_chisq[i] <- a$Chisq[2]
    lrt_df[i]    <- a$Df[2]
    raw_p[i]     <- a$`Pr(>Chisq)`[2]
  }

  # 5. Format p-values
  format_p <- function(p) {
    if (is.na(p)) return(NA_character_)
    if (p < 0.0001) return("< .0001")
    if (p < 0.001)  return("< .001")
    if (p < 0.01)   return(sprintf("= %.3f", p))
    return(sprintf("= %.3f", p))
  }

  LRT_p <- vapply(raw_p, format_p, FUN.VALUE = character(1))

  # 6. ΔAIC relative to previous model
  delta_prev <- c(NA, diff(aic_vals))

  tibble::tibble(
    Model           = names(model_list),
    Specification   = formulas,
    AIC             = aic_vals,
    Delta_AIC_prev  = delta_prev,
    R2_marginal     = r2_df$R2_marginal,
    R2_conditional  = r2_df$R2_conditional,
    LRT_chisq       = lrt_chisq,
    LRT_df          = lrt_df,
    LRT_p           = LRT_p
  )
}

```

```{r}
extract_spec_pretty <- function(model) {

  # 1. Get raw formula string without R's line breaks
  f_raw <- paste(deparse(model@call$formula, width.cutoff = 500), collapse = " ")

  # 2. Normalize spacing
  f_raw <- gsub("\\s+", " ", f_raw)

  # 3. Split into fixed and random parts
  #    This finds the first "(" that starts a random-effect term.
  split_pos <- regexpr("\\(1 \\|", f_raw)

  if (split_pos == -1) {
    # No random effects
    fixed <- f_raw
    random <- ""
  } else {
    fixed  <- trimws(substr(f_raw, 1, split_pos - 1))
    random <- trimws(substr(f_raw, split_pos, nchar(f_raw)))
  }

  # 4. Clean random effects:
  #    remove internal newlines, enforce single-line spacing
  random <- gsub("\\s*\\+\\s*", " + ", random)
  random <- gsub("\\s+", " ", random)

  # 5. Rebuild the final pretty spec:
  pretty <- paste0(fixed, "<br>", random)

  html(pretty)
}
```

## combined model: convergence ladder

Roles: `Ethnicity` = annotator trait, `spectrum` = annotator trait, `party` = speaker trait.

Interpret the highest rung that fits. Descend only on convergence failure, never on non-significance.

### rung 1

```{r}
# rung1_mod <- lmer(rating ~ value * Ethnicity * spectrum * party + (1|item_ID) + (1 | participant_ID),
#   data = wave_2_dfs$wave_2_speeches, REML = FALSE)
# saveRDS(rung1_mod, here(path, "wave_2_speeches_rung1_mod.RDS"))

# after first run, comment the fit/save and switch to readRDS
rung1_mod <- readRDS(here(path, "wave_2_speeches_rung1_mod.RDS"))

rung1_mod@optinfo$conv$lme4
```

```{r}
# On a convergence warning for rung1_mod, do NOT drop terms.
# Retry optimizers in this order first: allFit(), then bobyqa, then Nelder_Mead, then more iterations.

# rung1_allfit <- allFit(rung1_mod)
# summary(rung1_allfit)
#
# rung1_mod_bobyqa <- update(rung1_mod, control = lmerControl(optimizer = "bobyqa"))
# rung1_mod_nm     <- update(rung1_mod, control = lmerControl(optimizer = "Nelder_Mead"))
# rung1_mod_iter   <- update(rung1_mod, control = lmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5)))
```

### rung 2 (fallback — four-way interaction dropped)

```{r}
# rung2_mod <- lmer(rating ~ (value + Ethnicity + spectrum + party)^3 + (1|item_ID) + (1 | participant_ID),
#   data = wave_2_dfs$wave_2_speeches, REML = FALSE)
# saveRDS(rung2_mod, here(path, "wave_2_speeches_rung2_mod.RDS"))
#
rung2_mod <- readRDS(here(path, "wave_2_speeches_rung2_mod.RDS"))
rung2_mod@optinfo$conv$lme4
```

### rung 3 (fallback — Ethnicity:spectrum:party also dropped)

```{r}
# rung3_mod <- lmer(rating ~ value*Ethnicity*spectrum + value*Ethnicity*party + value*spectrum*party +
#     (1|item_ID) + (1 | participant_ID),
#   data = wave_2_dfs$wave_2_speeches, REML = FALSE)
# saveRDS(rung3_mod, here(path, "wave_2_speeches_rung3_mod.RDS"))
#
# rung3_mod <- readRDS(here(path, "wave_2_speeches_rung3_mod.RDS"))
# rung3_mod@optinfo$conv$lme4
```

### select rung

```{r}
# set this to whichever rung converged; everything below follows automatically
selected_mod <- rung2_mod   # rung2_mod / rung3_mod
selected_rung <- 2
```

### convergence record

```{r}
cat(paste0(
  "Rung used: ", selected_rung, "\n",
  "Formula: ", paste(deparse(formula(selected_mod)), collapse = " "), "\n",
  "Convergence warnings: [Andrew: fill in after running]\n",
  "Optimizers tried: [Andrew: fill in — none needed / allFit() / bobyqa / Nelder_Mead / more iterations]\n",
  "allFit() output: [Andrew: paste summary(allFit(...)) here if it was run]\n"
))
```

### model comparison

```{r}
# rung2_mod is the reporting row (three-way ceiling); rung1_mod is selected_mod's
# ML-fit twin (converged, interpreted model, unchanged). Each group tests every
# row against the row directly above it (model_comparison_table() does i-1 -> i,
# not first-row -> i), so the bottom row of each group is the rung2 -> rung1
# (four-way) test and is identical in both groups.

speeches_a_mods_final <- list(
  speeches1a_mod = speeches1a_mod,
  rung2_mod      = rung2_mod,
  rung1_mod      = rung1_mod)

speeches_b_mods_final <- list(
  speeches1b_mod = speeches1b_mod,
  rung2_mod      = rung2_mod,
  rung1_mod      = rung1_mod)

speech_tbls <- bind_rows(
  model_comparison_table(speeches_a_mods_final) |> mutate(Group = "Ethnicity Models"),
  model_comparison_table(speeches_b_mods_final) |> mutate(Group = "Spectrum Models")) |>

  mutate(Model = gsub("_mod", "", Model))

speech_tbl <- speech_tbls |> select(-Specification) |>

  gt(groupname_col = "Group") |>

  cols_label(
    Model           = "Model",
    AIC             = "AIC",
    Delta_AIC_prev  = html("$\\Delta$AIC"),
    R2_marginal     = html("$R^2_m$"),
    R2_conditional  = html("$R^2_c$"),
    LRT_chisq       = html("$\\chi^2$"),
    LRT_df          = html("df"),
    LRT_p           = html("p")
  ) |>

  fmt_number(
    columns = c(AIC, Delta_AIC_prev, R2_marginal, R2_conditional, LRT_chisq),
    decimals = 3) |>

  tab_header(title = "Model Comparisons: Ethnicity, and Political Spectrum") |>

  tab_source_note(source_note = "The rung1 row (four-way interaction test, rung2 vs rung1) is the same comparison in both groups, since rung1_mod and rung2_mod are shared across the Ethnicity and Spectrum ladders — the statistic is not independent between groups.")

speech_tbl
```

### model specification table

```{r}
speeches_a_specs <- tibble(
  Group = "Ethnicity Models",
  Model = names(speeches_a_mods_final),
  Specification = purrr::map(speeches_a_mods_final, extract_spec_pretty)
)

speeches_b_specs <- tibble(
  Group = "Political Spectrum Models",
  Model = names(speeches_b_mods_final),
  Specification = purrr::map(speeches_b_mods_final, extract_spec_pretty)
)

spec_table <- bind_rows(speeches_a_specs, speeches_b_specs) |> mutate(Model = gsub("_mod", "", Model))

gt_specs <- spec_table |>
  gt(groupname_col = "Group") |>
  cols_width(
    Model ~ px(150),
    Specification ~ px(400)
  ) |>
  cols_align("left", columns = Specification) |>
  tab_style(
    style = cell_text(weight = "bold", align = "center"),
    locations = cells_row_groups()
  ) |>
  tab_header(
    title = "Models Specifications: Ethnicity, and Political Spectrum"
  )

gt_specs
```

```{r}
gtsave(speech_tbl, here("tables", "main_speech_model_comparisons.tex"))
gtsave(gt_specs, here("tables", "main_speech_model_specifications.tex"))
```

### final model (REML = TRUE)

```{r}
selected_mod <- update(selected_mod, REML = TRUE)
# selected_rung noted above (selected_rung); rung 1 = full four-way interaction
saveRDS(selected_mod, here(path, "wave_2_speeches_combined_mod.RDS"))

# after first run, comment the fit/save and switch to readRDS
# selected_mod <- readRDS(here(path, "wave_2_speeches_combined_mod.RDS"))

anova <- anova(selected_mod, type = "III")
anova

eta_df <- eta_squared(selected_mod, partial = TRUE, ci = 0.95)
eta_df <- eta_df |>
  select(Parameter, Eta2_partial, CI_low, CI_high) |>
  rename(Effect = Parameter)
eta_df
```
