```{r include=FALSE}
library("here")         # file logistics
library("lme4")         # mixed effects metric
library("lmerTest")     # p values
library("tidyverse")    # code logistics
library("broom.mixed")  # model outputs as data frames
library("emmeans")      # estimated marginal means
library("patchwork")    # merge plots
library("report")       # apa reporting
library("gt")           # latex tables
library("effectsize")   # partial eta sq calculations

# center categorical predictors: contr.sum recodes to deviations from the average
options(scipen = 999, contrasts = c("contr.sum", "contr.poly"))
path <- here("_data", "_intermediary_data", "models")

# read in data
wave_2_dfs <- readRDS(here("_data", "_intermediary_data", "wave_2_dfs.RDS"))
```

### Appendix G

#### models in multiverse

Models:

 speeches_1a  pre-registered, ethnicity annotator trait
              rating ~ value + Ethnicity * party
 speeches_1b  pre-registered, spectrum annotator trait
              rating ~ value + spectrum * party
 speeches_2a  best-fitting, ethnicity annotator trait
              rating ~ value * Ethnicity * party
 speeches_2b  best-fitting, spectrum annotator trait
              rating ~ value * spectrum * party
 speeches_3a  combined model, all three-way interactions (SELECTED)
              rating ~ (value + Ethnicity + spectrum + party)^3
 speeches_3b  combined model, full four-way interaction (REJECTED — four-way adds no fit)
              rating ~ value * Ethnicity * spectrum * party

 Both combined models converge. speeches_3a is selected on parsimony: the
 four-way interaction in speeches_3b provides no improvement in fit
 (LRT p = .748, ΔAIC = +61.3 against speeches_3a).


```{r include=FALSE}
#### Pre-registered and best-fitting models

# --- speeches_1a: pre-registered, ethnicity ---
# speeches_1a <- lmer(
#   rating ~ value + Ethnicity * party + (1|item_ID) + (1|participant_ID),
#   data = wave_2_dfs$wave_2_speeches, REML = FALSE)
# saveRDS(speeches_1a, here(path, "wave_2_speeches_1a_mod.RDS"))
speeches_1a <- readRDS(here(path, "wave_2_speeches_1a_mod.RDS"))

# --- speeches_2a: value * ethnicity ---
# speeches_2a <- lmer(
#   rating ~ value * Ethnicity * party + (1|item_ID) + (1|participant_ID),
#   data = wave_2_dfs$wave_2_speeches, REML = FALSE)
# saveRDS(speeches_2a, here(path, "wave_2_speeches_2a_mod.RDS"))
speeches_2a <- readRDS(here(path, "wave_2_speeches_2a_mod.RDS"))

# --- speeches_1b: pre-registered, spectrum ---
# speeches_1b <- lmer(
#   rating ~ value + spectrum * party + (1|item_ID) + (1|participant_ID),
#   data = wave_2_dfs$wave_2_speeches, REML = FALSE)
# saveRDS(speeches_1b, here(path, "wave_2_speeches_1b_mod.RDS"))
speeches_1b <- readRDS(here(path, "wave_2_speeches_1b_mod.RDS"))

# --- speeches_2b: value * spectrum ---
# speeches_2b <- lmer(
#   rating ~ value * spectrum * party + (1|item_ID) + (1|participant_ID),
#   data = wave_2_dfs$wave_2_speeches, REML = FALSE)
# saveRDS(speeches_2b, here(path, "wave_2_speeches_2b_mod.RDS"))
speeches_2b <- readRDS(here(path, "wave_2_speeches_2b_mod.RDS"))
```


```{r include=FALSE}
## Combined models

# Roles: `Ethnicity` = annotator trait, `spectrum` = annotator trait, `party` = speaker trait.
# Both combined models converge cleanly. Selection rests on fit and parsimony (LRT/AIC),
# not convergence: speeches_3a is preferred because the four-way interaction in
# speeches_3b adds no improvement in fit.

# --- speeches_3b: full four-way interaction ---
# speeches_3b <- lmer(
#   rating ~ value * Ethnicity * spectrum * party + (1|item_ID) + (1|participant_ID),
#   data = wave_2_dfs$wave_2_speeches, REML = FALSE)
# saveRDS(speeches_3b, here(path, "wave_2_speeches_3b_mod.RDS"))
speeches_3b <- readRDS(here(path, "wave_2_speeches_3b_mod.RDS"))

# --- speeches_3a: all three-way interactions (SELECTED) ---
# speeches_3a <- lmer(
#   rating ~ (value + Ethnicity + spectrum + party)^3 + (1|item_ID) + (1|participant_ID),
#   data = wave_2_dfs$wave_2_speeches, REML = FALSE)
# saveRDS(speeches_3a, here(path, "wave_2_speeches_3a_mod.RDS"))
speeches_3a <- readRDS(here(path, "wave_2_speeches_3a_mod.RDS"))

# --- Indirection object ---
selected_mod  <- speeches_3a
selected_name <- "speeches_3a"
```


```{r include=FALSE}
# Nakagawa pseudo-R2 and AIC for a single model, as a one-row tibble.
model_fit_row <- function(model, name) {
  r2 <- as.data.frame(performance::r2_nakagawa(model))
  tibble::tibble(
    Model          = name,
    AIC            = AIC(model),
    R2_marginal    = r2$R2_marginal,
    R2_conditional = r2$R2_conditional)
}

# LRT of a larger model against a reference (nested) model.
# Returns chisq, df, and formatted p for the larger model.
lrt_row <- function(reference, larger) {
  a <- suppressWarnings(anova(reference, larger))
  p <- a$`Pr(>Chisq)`[2]
  p_fmt <- if (is.na(p)) NA_character_
           else if (p < 0.0001) "< .0001"
           else if (p < 0.001)  "< .001"
           else sprintf("= %.3f", p)
  tibble::tibble(
    LRT_chisq = a$Chisq[2],
    LRT_df    = a$Df[2],
    LRT_p     = p_fmt)
}
```

```{r include=FALSE}
extract_spec_pretty <- function(model) {
  f_raw     <- paste(deparse(model@call$formula, width.cutoff = 500), collapse = " ")
  f_raw     <- gsub("\\s+", " ", f_raw)
  split_pos <- regexpr("\\(1 \\|", f_raw)
  if (split_pos == -1) {
    fixed <- f_raw; random <- ""
  } else {
    fixed  <- trimws(substr(f_raw, 1, split_pos - 1))
    random <- trimws(substr(f_raw, split_pos, nchar(f_raw)))
  }
  random <- gsub("\\s*\\+\\s*", " + ", random)
  random <- gsub("\\s+", " ", random)
  html(paste0(fixed, "<br>", random))
}
```

#### Model comparison table

Three sections:

- **Ethnicity Models:** speeches_2a tested against speeches_1a (sequential LRT).
- **Spectrum Models:** speeches_2b tested against speeches_1b (sequential LRT).
- **Combined Model:** speeches_3a tested against BOTH best-fitting models
  (vs speeches_2a and vs speeches_2b — shown as separate rows, since these are
  different nested comparisons), then speeches_3b tested against speeches_3a to
  show the four-way interaction adds no fit.

ΔAIC in each row is relative to the model that row is tested against.

```{r}
# --- Ethnicity group: 1a -> 2a ---
eth_rows <- bind_rows(
  bind_cols(
    model_fit_row(speeches_1a, "speeches_1a"),
    tibble::tibble(Delta_AIC = NA_real_, LRT_chisq = NA_real_, LRT_df = NA_real_, LRT_p = NA_character_),
    tibble::tibble(Tested_against = NA_character_)),
  bind_cols(
    model_fit_row(speeches_2a, "speeches_2a"),
    tibble::tibble(Delta_AIC = AIC(speeches_2a) - AIC(speeches_1a)),
    lrt_row(speeches_1a, speeches_2a),
    tibble::tibble(Tested_against = "speeches_1a"))) |>
  mutate(Group = "Ethnicity Models")

# --- Spectrum group: 1b -> 2b ---
spec_rows <- bind_rows(
  bind_cols(
    model_fit_row(speeches_1b, "speeches_1b"),
    tibble::tibble(Delta_AIC = NA_real_, LRT_chisq = NA_real_, LRT_df = NA_real_, LRT_p = NA_character_),
    tibble::tibble(Tested_against = NA_character_)),
  bind_cols(
    model_fit_row(speeches_2b, "speeches_2b"),
    tibble::tibble(Delta_AIC = AIC(speeches_2b) - AIC(speeches_1b)),
    lrt_row(speeches_1b, speeches_2b),
    tibble::tibble(Tested_against = "speeches_1b"))) |>
  mutate(Group = "Spectrum Models")

# --- Combined section: 3a vs 2a, 3a vs 2b, 3b vs 3a ---
combined_rows <- bind_rows(
  bind_cols(
    model_fit_row(speeches_3a, "speeches_3a"),
    tibble::tibble(Delta_AIC = AIC(speeches_3a) - AIC(speeches_2a)),
    lrt_row(speeches_2a, speeches_3a),
    tibble::tibble(Tested_against = "speeches_2a")),
  bind_cols(
    model_fit_row(speeches_3a, "speeches_3a"),
    tibble::tibble(Delta_AIC = AIC(speeches_3a) - AIC(speeches_2b)),
    lrt_row(speeches_2b, speeches_3a),
    tibble::tibble(Tested_against = "speeches_2b")),
  bind_cols(
    model_fit_row(speeches_3b, "speeches_3b"),
    tibble::tibble(Delta_AIC = AIC(speeches_3b) - AIC(speeches_3a)),
    lrt_row(speeches_3a, speeches_3b),
    tibble::tibble(Tested_against = "speeches_3a"))) |>
  mutate(Group = "Combined Model")

comparison_data <- bind_rows(eth_rows, spec_rows, combined_rows) |>
  select(Group, Model, Tested_against, AIC, Delta_AIC,
         R2_marginal, R2_conditional, LRT_chisq, LRT_df, LRT_p)

speech_tbl <- comparison_data |>
  gt(groupname_col = "Group") |>
  cols_label(
    Model          = "Model",
    Tested_against = "vs.",
    AIC            = "AIC",
    Delta_AIC      = html("$\\Delta$AIC"),
    R2_marginal    = html("$R^2_m$"),
    R2_conditional = html("$R^2_c$"),
    LRT_chisq      = html("$\\chi^2$"),
    LRT_df         = html("df"),
    LRT_p          = html("p")) |>
  fmt_number(
    columns = c(AIC, Delta_AIC, R2_marginal, R2_conditional, LRT_chisq),
    decimals = 3) |>
  sub_missing(columns = everything(), missing_text = "—") |>
  tab_style(
    style = cell_text(align = "center"),
    locations = cells_row_groups()) |>
  tab_header(
    title    = "Model Comparisons: Ethnicity and Political Spectrum",
    subtitle = "Best-fitting and combined models tested by likelihood ratio against nested references") |>
  tab_source_note(
    source_note = paste0(
      "χ², df, and p are likelihood ratio tests of each model against the model named in the ",
      "'vs.' column; ΔAIC is relative to that same reference. speeches_3a is shown twice in the ",
      "Combined section because it is tested against both best-fitting models (speeches_2a and ",
      "speeches_2b); it improves on both. speeches_3b (full four-way interaction) adds no fit over ",
      "speeches_3a, so speeches_3a is selected on parsimony. Both combined models converge."))

speech_tbl
gtsave(speech_tbl, here("tables", "main_speech_model_comparisons.tex"))
```

#### Model specification table

```{r}
all_specs <- bind_rows(
  tibble(
    Group = "Ethnicity Models",
    Model = c("speeches_1a", "speeches_2a"),
    Specification = purrr::map(list(speeches_1a, speeches_2a), extract_spec_pretty)),
  tibble(
    Group = "Spectrum Models",
    Model = c("speeches_1b", "speeches_2b"),
    Specification = purrr::map(list(speeches_1b, speeches_2b), extract_spec_pretty)),
  tibble(
    Group = "Combined Models",
    Model = c("speeches_3a", "speeches_3b"),
    Specification = purrr::map(list(speeches_3a, speeches_3b), extract_spec_pretty)))

gt_specs <- all_specs |>
  gt(groupname_col = "Group") |>
  cols_width(
    Model ~ px(220),
    Specification ~ px(400)) |>
  cols_align("left", columns = Specification) |>
  tab_style(
    style = cell_text(weight = "bold", align = "center"),
    locations = cells_row_groups()) |>
  tab_header(title = "Model Specifications")

gt_specs
gtsave(gt_specs, here("tables", "main_speech_model_specifications.tex"))
```

#### Final model Parameter Estimates

```{r}
# Refit the selected model with REML = TRUE for estimation, and save the ANOVA
# and partial eta-squared. These three objects are read by the analysis notebook
# (5_speech_model_analysis) under exactly these filenames.

#selected_mod <- update(selected_mod, REML = TRUE)
#saveRDS(selected_mod, here(path, "wave_2_speeches_mod_3b_REML.RDS"))
selected_mod <- readRDS(here(path, "wave_2_speeches_mod_3b_REML.RDS"))

# anova_selected <- anova(selected_mod, type = "III")
# saveRDS(anova_selected, here(path, "wave_2_speeches_mod_3b_REML_anova.RDS"))
anova_selected <- readRDS(here(path, "wave_2_speeches_mod_3b_REML_anova.RDS"))
anova_selected

# eta_df <- eta_squared(selected_mod, partial = TRUE, ci = 0.95)
# eta_df <- eta_df |>
#   select(Parameter, Eta2_partial, CI_low, CI_high) |>
#   rename(Effect = Parameter)
# saveRDS(eta_df, here(path, "wave_2_speeches_mod_3b_REML_eta.RDS"))
eta_df <- readRDS(here(path, "wave_2_speeches_mod_3b_REML_eta.RDS"))
eta_df
```
