

```{r include=FALSE}
library("here")         # file logistics
library("lme4")         # mixed effects metric
library("lmerTest")     # p values
library("tidyverse")    # code logistics
library("emmeans")      # estimated marginal means
library("patchwork")    # merge plots
library("gt")           # latex tables
library("effectsize")   # partial eta sq calculations

options(scipen = 999, contrasts = c("contr.sum", "contr.poly"))
path <- here("_data", "_intermediary_data", "models")

wave_2_dfs <- readRDS(here("_data", "_intermediary_data", "wave_2_dfs.RDS"))
speeches_dfs <- wave_2_dfs[["wave_2_speeches"]]

# Selected model (rung-2 combined, REML = TRUE)
selected_mod <- readRDS(here(path, "wave_2_speeches_mod_3b_REML.RDS"))

# Total SD across all variance components — used for Cohen's d throughout
total_sd <- sqrt(sum(sapply(lme4::VarCorr(selected_mod), function(v) sum(diag(as.matrix(v))))))
```

### Results

#### table: subgroup deviation from overall sample

```{r}
# Mean rating per speech x value, for overall sample
overall_item_profile <- speeches_dfs |>
  filter(!is.na(rating)) |>
  group_by(item_ID, value) |>
  summarise(overall_mean = mean(rating, na.rm = TRUE), .groups = "drop")

# Mean rating per speech x value, per subgroup
subgroup_item_profiles <- speeches_dfs |>
  filter(!is.na(rating)) |>
  group_by(Ethnicity, spectrum, item_ID, value) |>
  summarise(subgroup_mean = mean(rating, na.rm = TRUE), .groups = "drop")

# Join and correlate across item x value combinations
subgroup_summary <- subgroup_item_profiles |>
  left_join(overall_item_profile, by = c("item_ID", "value")) |>
  group_by(Ethnicity, spectrum) |>
  summarise(
    n_obs      = n(),
    MAD        = mean(abs(subgroup_mean - overall_mean)),
    r_pearson  = cor(subgroup_mean, overall_mean),
    r_spearman = cor(subgroup_mean, overall_mean, method = "spearman"),
    .groups    = "drop"
  ) |>
  arrange(desc(MAD))

gt_subgroup <- subgroup_summary |>
  gt() |>
  cols_label(
    Ethnicity  = "Ethnicity",
    spectrum   = "Political Spectrum",
    n_obs      = md("*n* cells"),
    MAD        = md("MAD"),
    r_pearson  = md("*r* (Pearson)"),
    r_spearman = md("*r* (Spearman)")
  ) |>
  fmt_number(columns = c(MAD, r_pearson, r_spearman), decimals = 3) |>
  tab_header(
    title    = "Subgroup Deviation from Overall Sample Value Profiles",
    subtitle = "Mean Absolute Deviation and profile correlations per Ethnicity × Political Spectrum subgroup"
  ) |>
  tab_footnote(
    footnote = "MAD = mean absolute deviation from overall sample mean per value (rating scale units).
                Pearson r and Spearman ρ reflect agreement between subgroup and overall value profiles.",
    locations = cells_column_labels(columns = MAD)
  ) |>
  cols_align("center", columns = c(n_obs, MAD, r_pearson, r_spearman)) |>
  tab_options(table.font.size = 12, data_row.padding = px(4))

gt_subgroup
gtsave(gt_subgroup, here("tables", "wave_2_speeches_subgroup_deviation.tex"))
```


```{r include=FALSE}
build_anova_gt <- function(anova_obj, eta_df, title = "Type III ANOVA Results") {

  anova_df <- anova_obj |>
    as.data.frame() |>
    tibble::rownames_to_column("Effect") |>
    mutate(Effect = trimws(Effect))

  eta_df <- eta_df |>
    select(Effect, Eta2_partial) |>
    mutate(Effect = trimws(Effect)) |>
    distinct()

  df <- left_join(anova_df, eta_df, by = "Effect") |>
    mutate(
      p_fmt = case_when(
        is.na(`Pr(>F)`)  ~ NA_character_,
        `Pr(>F)` < .0001 ~ "p < .0001",
        `Pr(>F)` < .001  ~ "p < .001",
        TRUE             ~ sprintf("p = %.3f", `Pr(>F)`)
      ),
      F_fmt   = sprintf("F(%d, %s) = %.2f", NumDF,
                        format(round(DenDF), scientific = FALSE), `F value`),
      eta_fmt = sprintf("%.3f", Eta2_partial),
      Group = case_when(
        !str_detect(Effect, ":") ~ "Main Effects",
        str_count(Effect, ":") == 1 ~ "Two-Way Interactions",
        str_count(Effect, ":") == 2 ~ "Three-Way Interactions",
        TRUE ~ "Four-Way Interactions"
      )
    )

  df |>
    select(Group, Effect, F_fmt, p_fmt, eta_fmt) |>
    gt(groupname_col = "Group") |>
    cols_label(
      Effect  = "Effect",
      F_fmt   = md("$F$"),
      p_fmt   = md("$p$"),
      eta_fmt = md("$\\eta^2_{partial}$")
    ) |>
    cols_align("center", columns = c(F_fmt, p_fmt, eta_fmt)) |>
    tab_style(style = cell_text(align = "center"), locations = cells_row_groups()) |>
    tab_header(title = title)
}
```

### Appendix G

#### Table 10 — Pre-registered models (speeches1a, speeches1b)

Type III ANOVAs for the two pre-registered models. These feed the hand-merged
Table 10 comparator (`main_speeches_mod_1a_1b_anova.tex`).

```{r}
# Pre-registered models (ethnicity and spectrum) — feed the Table 10 merge
speeches_a_mods <- readRDS(here(path, "speeches_a_mods.RDS"))
speeches_b_mods <- readRDS(here(path, "speeches_b_mods.RDS"))

speeches1a_mod <- speeches_a_mods[["speeches1a_mod"]]
anova_1a <- anova(speeches1a_mod)
eta_1a <- eta_squared(speeches1a_mod, partial = TRUE, ci = 0.95) |>
  select(Parameter, Eta2_partial, CI_low, CI_high) |>
  rename(Effect = Parameter)
tbl_1a <- build_anova_gt(anova_1a, eta_1a, title = "Type III ANOVA: speeches_1a (pre-registered, ethnicity)")
tbl_1a

gtsave(tbl_1a, here("tables", "wave_2_speeches_mod_1a_anova.tex"))
```

```{r}
speeches1b_mod <- speeches_b_mods[["speeches1b_mod"]]
anova_1b <- anova(speeches1b_mod)
eta_1b <- eta_squared(speeches1b_mod, partial = TRUE, ci = 0.95) |>
  select(Parameter, Eta2_partial, CI_low, CI_high) |>
  rename(Effect = Parameter)
tbl_1b <- build_anova_gt(anova_1b, eta_1b, title = "Type III ANOVA: speeches_1b (pre-registered, spectrum)")
tbl_1b
gtsave(tbl_1b, here("tables", "wave_2_speeches_mod_1b_anova.tex"))
```

#### Table 7 — Speeches ANOVA (selected combined model)

Type III ANOVA and partial η² for the selected rung-2 combined model (speeches_3a).
Reads cached objects from the model-selection notebook.

```{r}
anova_selected <- readRDS(here(path, "wave_2_speeches_combined_mod_anova.RDS"))
eta_selected   <- readRDS(here(path, "wave_2_speeches_combined_mod_eta.RDS"))

tbl_selected <- build_anova_gt(anova_selected, eta_selected,
                               title = "Type III ANOVA: speeches_3a (selected combined model)")
tbl_selected
gtsave(tbl_selected, here("tables", "wave_2_speeches_mod_4_anova.tex"))
```

```{r include=FALSE}
labeller_value <- labeller(value = function(x) gsub("value: ", "", x))
```

#### in-text: speeches model main effects

```{r include=FALSE}
# APA sentence helper: pairwise contrasts with z-tests and Cohen's d
apa_emm_pairs <- function(model, specs, total_sd, digits = 2) {

  cmp <- emmeans::emmeans(model, specs, lmer.df = "asymptotic") |>
    emmeans::contrast("pairwise", adjust = "tukey") |>
    broom::tidy()

  cmp |>
    dplyr::mutate(
      est   = sprintf(paste0("%.", digits, "f"), estimate),
      se    = sprintf(paste0("%.", digits, "f"), std.error),
      z_fmt = sprintf(paste0("%.", digits, "f"), statistic),
      p     = dplyr::case_when(
        adj.p.value < .001 ~ "< .001",
        TRUE ~ sprintf("= %.3f", adj.p.value)),
      d_fmt = sprintf(paste0("%.", digits, "f"), estimate / total_sd),
      apa   = sprintf(
        "%s: difference = %s, SE = %s, z = %s, p %s, d = %s.",
        contrast, est, se, z_fmt, p, d_fmt)) |>
    dplyr::pull(apa)
}
```

```{r}
apa_emm_pairs(selected_mod, specs = ~ Ethnicity, total_sd = total_sd)
```

```{r}
apa_emm_pairs(selected_mod, specs = ~ spectrum, total_sd = total_sd)
```

#### Figure 13: EMM by ethnicity and political spectrum

Two panels: ethnicity contrasts within spectrum (A), spectrum contrasts within
ethnicity (B). *(Chunk currently disabled — re-enable to produce the figure.)*

```{r}
p_A <- plot(emmeans(selected_mod, ~ Ethnicity | spectrum), comparisons = TRUE) +
  facet_wrap(~spectrum, ncol = 1) +
  ylab("Ethnicity") + xlab("Estimated Marginal Mean") +
  theme_classic()

p_B <- plot(emmeans(selected_mod, ~ spectrum | Ethnicity), comparisons = TRUE) +
  facet_wrap(~Ethnicity, ncol = 1) +
  ylab("Political Spectrum") + xlab("Estimated Marginal Mean") +
  theme_classic()

(p_A | p_B) + plot_annotation(tag_levels = "A")
# ggsave(here("images", "main_emmeans_speeches_by_ethnicity_spectrum.pdf"))
```

#### Figure 15: EMM by ethnicity by political spectrum

```{r}
emm <- emmeans(selected_mod, ~ spectrum | Ethnicity, lmer.df = "asymptotic")
pdat <- plot(emm, comparisons = TRUE, plotit = FALSE)

pdat_l <- pdat |> filter(!is.na(lcmpl))
pdat_r <- pdat |> filter(!is.na(rcmpl))

ggplot(pdat, aes(x = the.emmean, y = spectrum)) +
  geom_rect(aes(xmin = asymp.LCL, xmax = asymp.UCL,
                ymin = as.numeric(spectrum) - 0.5,
                ymax = as.numeric(spectrum) + 0.5,
                fill = Ethnicity),
            alpha = 0.6) +
  scale_fill_viridis_d(option = "D") +
  geom_segment(data = pdat_l,
               aes(xend = lcmpl, yend = spectrum),
               arrow = arrow(length = unit(0.15, "cm")), colour = "red") +
  geom_segment(data = pdat_r,
               aes(xend = rcmpl, yend = spectrum),
               arrow = arrow(length = unit(0.15, "cm")), colour = "red") +
  geom_point() +
  coord_cartesian(xlim = c(2, 5)) +
  facet_wrap(~ Ethnicity, ncol = 1) +
  theme_classic() + xlab("Estimated Marginal Mean") + ylab("Political Spectrum") +
  theme(legend.position = "none")

ggsave(here("images", "main_emmeans_spectrum_ethnicity_by_party.pdf"))
```



```{r include=FALSE}
### Supporting — ethnicity × value × spectrum contrasts


# cmp <- emmeans::emmeans(
#   selected_mod,
#   specs    = ~ Ethnicity | value * spectrum,
#   lmer.df  = "asymptotic") |>
#   emmeans::contrast("pairwise", adjust = "tukey") |>
#   broom::tidy() |>
#   dplyr::filter(adj.p.value < .05) |>
#   dplyr::mutate(
#     SE          = std.error,
#     adj.p.value = dplyr::case_when(
#       adj.p.value < .0001 ~ "< .0001",
#       adj.p.value < .001  ~ "< .001",
#       TRUE ~ sprintf("= %.3f", adj.p.value)))
#
# put_black_first <- function(cmp_df, black_label = "(Black/African American)") {
#   cmp_df |>
#     dplyr::group_by(value, spectrum) |>
#     tidyr::separate(contrast, into = c("g1", "g2"), sep = " - ", remove = FALSE) |>
#     dplyr::mutate(
#       flip      = g2 == black_label,
#       g1_new    = dplyr::if_else(flip, g2, g1),
#       g2_new    = dplyr::if_else(flip, g1, g2),
#       estimate  = dplyr::if_else(flip, -estimate, estimate),
#       statistic = dplyr::if_else(flip, -statistic, statistic),
#       contrast  = paste0(g1_new, " - ", g2_new)) |>
#     dplyr::ungroup() |>
#     dplyr::select(-g1, -g2, -g1_new, -g2_new, -flip)
# }
#
# cmp_fixed <- put_black_first(cmp)
#
# contrasts_gt <- function(df, title = "Pairwise Comparisons") {
#   df |>
#     dplyr::mutate(
#       contrast = contrast |>
#         stringr::str_replace_all("\\(", "") |>
#         stringr::str_replace_all("\\)", "")) |>
#     dplyr::transmute(
#       Value     = value,
#       Spectrum  = spectrum,
#       Contrast  = contrast,
#       Estimate  = round(estimate, 2),
#       SE        = round(SE, 2),
#       Statistic = sprintf("z = %.2f", round(statistic, 2)),
#       `p-value` = adj.p.value) |>
#     gt::gt() |>
#     gt::tab_header(title = title) |>
#     gt::cols_align(align = "center", columns = c(Estimate, SE, Statistic, `p-value`)) |>
#     gt::tab_options(table.font.size = 12, data_row.padding = px(4))
# }
#
# gt_contrast_table <- contrasts_gt(cmp_fixed,
#                                   title = "Significant Contrasts: Ethnicity and Political Leaning by Value")
# gt_contrast_table
# gtsave(gt_contrast_table, here("tables", "wave_2_speeches_ethnicity_value_contrasts.tex"))
```
