## Pilot Studies

```{r pilot_participants_setup, include=FALSE}
library("here")
library("tidyverse")
library("rlang")
library("gt")

data_path <- here("_data", "_participant_scores", "wave_2")

# read in data
pilot_2_dfs <- readRDS(here(data_path, "pilot_2_dfs.RDS"))

# rename ethnicity column 
pilot_2_dfs[['speeches2']] <- pilot_2_dfs[['speeches2']] |> 
  rename(Ethnicity = `Ethnicity simplified`)

# subset each df
pilot_2_demog_dfs <- lapply(pilot_2_dfs, function(df){
  df <- df |> select(participant_ID, Age, Sex, `Ethnicity`) |> unique()})

#re-add spectrum to speeches2 df
pilot_2_demog_dfs[['speeches2']] <- merge(
  
  pilot_2_demog_dfs[['speeches2']],
  
  pilot_2_dfs[['speeches2']] |> 
    select(participant_ID, `Political spectrum (us)`) |> 
    unique(), 
  
  by = "participant_ID", all.x=T)

rm(pilot_2_dfs)


# convert age to numeric
pilot_2_demog_dfs <- lapply(pilot_2_demog_dfs, function(df){df|>mutate(Age = as.numeric(Age))})

# recode Asian ethnicity
pilot_2_demog_dfs[["lyrics"]] <- pilot_2_demog_dfs[["lyrics"]] |> 
  mutate(Ethnicity = case_when(
    Ethnicity == "East Asian" | Ethnicity == "South East Asian" | Ethnicity == "South Asian" ~ "Asian", 
    .default = Ethnicity))

pilot_2_demog_dfs[["speeches1"]]<- pilot_2_demog_dfs[["speeches1"]] |> 
  mutate(Ethnicity = case_when(
    Ethnicity == "East Asian" | Ethnicity == "South East Asian" | Ethnicity == "South Asian" ~ "Asian", 
    .default = Ethnicity))

pilot_2_demog_dfs[["speeches2"]] <- pilot_2_demog_dfs[["speeches2"]] |> mutate(Ethnicity = case_when(
  Ethnicity == "DATA_EXPIRED" ~ NA, 
  .default = Ethnicity
))

# study labels
pilot_labels <- c(
  lyrics     = "Pilot 1 (Lyrics)",
  speeches1  = "Pilot 2 (Speeches: Ethnicity)",
  speeches2  = "Pilot 3 (Speeches: Political)")
```

```{r pilot_participants_table_functions, include=FALSE}
format_apa_table <- function(gt_tbl, title, subtitle = NULL) {
  gt_tbl %>%
    tab_header(
      title = title, 
      subtitle = subtitle
    ) %>%
    tab_options(
      table.font.names                  = c("Times New Roman", "Times", "serif"),
      table.border.top.width            = 1,  # top line
      column_labels.border.bottom.width = 1,  # header line
      table.border.bottom.width         = 1,  # bottom line
      table_body.hlines.width           = 0,  # no row lines
      data_row.padding                  = px(2)
    )
}

summarize_demographics <- function(df, study_name) {
  df %>%
    summarise(
      Study   = study_name,
      N       = n(),
      Age_M   = mean(Age, na.rm = TRUE),
      Age_SD  = sd(Age, na.rm = TRUE),
      Age_Min  = min(Age, na.rm = TRUE),
      Age_Max  = max(Age, na.rm = TRUE),
      Female_p  = mean(Sex == "Female", na.rm = TRUE) * 100) %>%
  mutate(
    `Age M (SD)` = sprintf("%.1f (%.1f)", Age_M, Age_SD),
    `Age range`  = sprintf("%d–%d", round(Age_Min), round(Age_Max)),
    `% Female`   = round(Female_p, 1)) %>%
  select(Study, N, `Age M (SD)`, `Age range`, `% Female`)
}

summarize_ethnicity <- function(df, study_name) {
  df %>%
    count(Ethnicity) %>%
    mutate(
      pct = 100 * n / sum(n),
      value = paste0(n, " (", round(pct, 1), "%)")
    ) %>%
    select(Ethnicity, value) %>%
    pivot_wider(names_from = Ethnicity, values_from = value) %>%
    mutate(Study = study_name, .before = 1)
}


summarize_spectrum <- function(df, study_name) {
  df %>%
    count(`Political spectrum (us)`) %>%
    mutate(
      pct = 100 * n / sum(n),
      value = paste0(n, " (", round(pct, 1), "%)")
    ) %>%
    select(`Political spectrum (us)`, value) %>%
    pivot_wider(names_from = `Political spectrum (us)`, values_from = value) %>%
    mutate(Study = study_name, .before = 1)
}
```

The primary aim of our pilot studies was to estimate the number of ratings needed per stimulus. We estimated the target number of ratings separately per text type (`lyrics` / `speeches`) as we expected that each form of text could vary in terms of degree of ambiguity. To do so, we followed a procedure similar to @DeBruine_Jones_2018. This process essentially involves gathering data comprised of a) a relatively large number of ratings, on b) a relatively small sub-sample of the items to be rated, followed by bootstrapping estimates of inter-rater reliability on a range of different possible numbers of ratings. Doing so shows a sampling distribution of inter-rater reliability estimates for e.g. 5 vs. 10 vs. 15 ratings per stimulus, allowing for an estimate of a target number of ratings to achieve a target inter-rater reliability. We bootstrapped estimates of inter-rater reliability by incrementally sub-sampling numbers of raters, 50 times per increment, and then estimating a) the Pearson correlation between the sub-sample increment mean and the overall sample mean, and b) inter-rater reliability [@DeBruine_Jones_2018]. We sub-sampled in increments of 5, ranging from 5 to 50, with replacement, 50 times for each increment: i.e. we sampled 5 raters 50 times, 10 raters 50 times, 15 raters 50 times etc. up to 50 raters 50 times, all with replacement. 

As we aim to represent the mean of a population of raters, we estimate inter-rater reliability as 'general consensus' (as opposed to 'absolute agreement') via intraclass correlation coefficient ICC(2,k)[@koo2016guideline]. The '2' indicates two-way random effects, in that both the targets of the ratings and the raters are treated as random samples from a larger population, and the 'k' indicates the reliabiliity of the average of *k* raters. In essence, ICC2k estimates how consistently a group mean rating of k raters reflects the "true" score of each item. We used a value of 0.75 as a threshold to indicate sufficent, or "good", agreement [@koo2016guideline]. Thus, for a given number of raters, we estimate a sampling distribution representing their degree of general consensus, and whether the the mean consensus exceeds the 0.75 threshold. Although we did not see a recommended threshold for pearson correlations in [@DeBruine_Jones_2018], we set a threshold of 0.9 to indicate sufficient, or "high" correlation between sub-sample and overall sample means.

### Participants

We recruited a total of 300 participants: 100 each stratified by ethnicity completed the `lyrics` and the first `speeches` pilot, and 100 participants stratified by political orientation completed the second `speeches` pilot. 1 participant each in the pilot study using lyrics and the first using speeches withdrew their data at the time this manuscript was written, as did 2 participants in the second pilot using speeches (see Table 1). The pilot study using lyrics and the first using speeches were stratified by Ethnicity (see Tables 2a), and the second pilot using speeches was stratified by Political Spectrum (see Tables 2b and Table 3).

```{r pilot_demogs}
# demographics table

pilot_demog_tbl <- imap_dfr(
  pilot_2_demog_dfs,
  ~ summarize_demographics(.x, pilot_labels[[.y]] %||% .y))

pilot_demog_tbl %>%
  gt() %>%
  cols_align(align = "left",  columns = c(Study)) %>%
  cols_align(align = "center", columns = c(N, `% Female`, `Age range`, `Age M (SD)`)) %>%
  fmt_number(columns = `% Female`, decimals = 1) %>%
  format_apa_table("Table 1", "Participant demographics — Pilot studies")
```



```{r}
pilot_ethnicity_tbl <- imap_dfr(
  pilot_2_demog_dfs[1:2], 
  ~ summarize_ethnicity(.x, pilot_labels[[.y]] %||% .y)) 

pilot_ethnicity_tbl |>
  gt() %>%
  cols_label(
    `Black/African American` = html("Black/<br>African American"),
    `Latino/Hispanic`        = html("Latino/<br>Hispanic"),
    Asian                    = "Asian",
    White                    = "White") |>
  
  fmt_number(columns = -Study, decimals = 1) %>%
  
  cols_align(align = "center", 
             columns = c(Asian, `Black/African American`, 
                         `Latino/Hispanic`, White)) %>%
  
  format_apa_table("Table 2a", "Participant ethnicity — Pilot studies")
```


```{r}
pilot_2_demog_dfs[[3]] |> 
    count(Ethnicity) %>%
    mutate(
      pct = 100 * n / sum(n),
      value = paste0(n, " (", round(pct, 1), "%)")) |> 
  select(Ethnicity, value) |>
  
  gt() |>
  
  cols_label(Ethnicity = "Ethnicity (Simplified)", value = "N (%)") |>
  cols_align(align = "center", columns = c(Ethnicity, value)) %>%
  format_apa_table("Table 2b", "Participant ethnicity, Pilot 3 Speeches: Political Spectrum")
```

```{r}
pilot_spectrum_tbl <- imap_dfr(
  pilot_2_demog_dfs["speeches2"],
  ~ summarize_spectrum(.x, pilot_labels[[.y]] %||% .y))

pilot_spectrum_tbl |>
  gt() %>%
  
  fmt_number(
    columns = -Study, 
    decimals = 1
  ) %>%
  
  cols_align(align = "center", columns = c(Conservative, Liberal, Moderate)) %>%
  
  format_apa_table("Table 3", "Participants by political leaning — Main studies")
```



