# Appendix A: Stimuli

```{r setup, include=FALSE}
library(tidyverse)
library(here)
library(viridis)
library(cowplot)
library(gt)
library(rlang)

speech_df <- readRDS(here("_data", "_primary_data", "speech_df.RDS"))
wave_2_dfs <- readRDS(here("_data", "_participant_scores", "wave_2", "wave_2_dfs.RDS"))
pilot_2_dfs <- readRDS(here("_data", "_participant_scores", "wave_2", "pilot_2_dfs.RDS"))

speech_IDs <- bind_rows(
    wave_2_dfs[["speeches1"]] |> select(item_ID), 
    wave_2_dfs[["speeches2"]] |> select(item_ID), 
    pilot_2_dfs[["speeches1"]] |> select(item_ID),
    pilot_2_dfs[["speeches2"]] |> select(item_ID)) |> 
  filter(!item_ID == "attention") |> unique() |> pull()

lyrics_IDs <- bind_rows(
    wave_2_dfs[["lyrics1"]] |> select(item_ID), 
    wave_2_dfs[["lyrics2"]] |> select(item_ID), 
    pilot_2_dfs[["lyrics"]] |> select(item_ID)) |> 
  filter(!item_ID == "attention") |> unique() |> pull()

lyrics_df <- readRDS(file = here("_data", "_primary_data", "lyrics_df.RDS"))
speech_df <- readRDS(here("_data", "_primary_data", "speech_df.RDS"))
```

```{r plot_function, include=FALSE}
colors <- viridis(10, option="plasma") 

# apa formatting
theme_apa <- function(base_size = 12, base_family = "sans") {
  theme_minimal(base_size = base_size, base_family = base_family) +
    theme(
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      axis.line = element_line(color = "black"),
      axis.ticks = element_line(color = "black"),
      axis.title = element_text(size = base_size, face = "plain"),
      axis.text = element_text(size = base_size * 0.9),
      #legend.position = "none",
      legend.title = element_text(face = "plain"))}


make_removed_counts <- function(df, col) {
  
  df |>  
    # remove songs that were not evaluated
    filter(selected %in% c("keep","remove")) |>
    
    # count the number of rows by `selected`
    count({{ col }}, selected, name = "n") |>
    
    # create a new column to order the factor
    group_by({{ col }}) |>
    mutate(removed_n = sum(n * (selected == "keep"))) |>
    ungroup() |>
    
    # order the factor by the number of rows in the given subgroup
    mutate(
      selected = factor(selected, levels = c("remove","keep")),
      # coerce to character so fct_reorder works for numeric columns too
      {{ col }} := fct_reorder(as.character({{ col }}), removed_n, .desc = FALSE)) |>
    
    # stacked column plot
    ggplot(aes(x = {{ col }}, y = n, fill = selected)) +
    geom_col(alpha = .75) +
    coord_flip() +
    #scale_fill_manual(values = c(colors[5], colors[9])) + 
    labs(x = NULL, y = "Count") +
    
    theme_apa()
}
```

```{r table_function, include=FALSE}
make_apa_freq_table <- function(df, bin_col,
                                title = "Table",
                                subtitle = "Frequencies by bin") {

  bin_sym   <- enquo(bin_col)
  bin_label <- as_label(bin_sym)   # for column header

  # 1) counts within levels of factor
  sum_df <- df %>%
    filter(selected %in% c("keep","remove")) %>%
    mutate(selected = factor(selected, levels = c("remove","keep"))) %>%
    count(!!bin_sym, selected, name = "n") %>%
    # ensure both levels exist for every bin
    tidyr::complete(!!bin_sym, selected, fill = list(n = 0)) %>%
    group_by(!!bin_sym) %>%
    summarise(
      Total        = sum(n),
      Removed_n    = sum(n[selected == "remove"]),
      Kept_n       = sum(n[selected == "keep"]),
      Removed_pct  = Removed_n / Total,
      Kept_pct     = Kept_n / Total,
      .groups = "drop"
    ) %>%
    arrange(desc(Removed_n)) %>%
    relocate(Total, .after = last_col())

  # 2) build gt table
  # Note: cols_merge() will merge the *_n and *_pct columns into a single display column.
  gt_tab <- sum_df %>%
    gt() %>%
    
    # define title and subtitle
    tab_header(title = title, subtitle = subtitle) %>%
    
    # center column headers
    tab_style(
      style = cell_text(align = "center"),
      locations = cells_column_labels(columns = everything())) |>
    
    # headers of each column
    cols_label(
      !!bin_sym   := bin_label,
      Total       = html("Total n"),
      Removed_n   = "Removed",
      Removed_pct = "Removed %",
      Kept_n      = "Kept",
      Kept_pct    = "Kept %"
    ) %>%
    
    # set number of decimal places
    fmt_number(columns = Total, decimals = 0) %>%
    fmt_number(columns = c(Removed_n, Kept_n), decimals = 0) %>%
    fmt_percent(columns = c(Removed_pct, Kept_pct), decimals = 1) %>%
    
    # merge n and % into "n (xx.x%)"
    cols_merge(
      columns = c(Removed_n, Removed_pct),
      pattern = "{1} ({2})") %>%
    cols_merge(
      columns = c(Kept_n, Kept_pct),
      pattern = "{1} ({2})") %>%
    cols_hide(columns = c(Removed_pct, Kept_pct)) %>%
    
    
    # set which lines show
    tab_options(
      table.font.size = 12,
      heading.align = "left",    
      
      table.border.top.width = px(0),
      table.border.bottom.width = px(0),
      
      # bold under the heading (title/subtitle)
      heading.border.bottom.width = px(2),
      
      # remove body row lines
      table_body.hlines.width = px(0),
      table_body.vlines.width = px(0),
      table_body.border.top.width = px(0),
      table_body.border.bottom.width = px(0),
      
      column_labels.border.top.width = px(0),
      column_labels.border.bottom.width = px(1)
    )

  gt_tab
}

```

```{r genre_function, include=FALSE}
make_genre_comparison_table <- function(df,
                                        title = "Table X",
                                        subtitle = "Top tf-ifd Terms of sample") {

  df %>%
    select(lda_genre_terms, artist_topic_bin_labels) %>%
    distinct() %>%
    arrange(lda_genre_terms) %>%
    gt() %>%
    tab_header(
      title = title,
      subtitle = subtitle
    ) %>%
    cols_label(
      lda_genre_terms         = "Top LDA terms of initial pool",
      artist_topic_bin_labels = "Top tf-ifd terms of sample"
    ) %>%
    tab_options(
      #table.font.names = c("Times New Roman", "Times", "serif"),
      table.font.size = 12,
      heading.align = "left",    
      # remove all body lines
      table_body.hlines.width = px(0),
      table_body.border.top.width = px(0),
      table_body.border.bottom.width = px(0),
      # header: no line above, single line below
      column_labels.border.top.width = px(0),
      column_labels.border.bottom.width = px(1),
      # remove outer box
      table.border.top.width = px(0),
      table.border.bottom.width = px(0),
      table.border.left.width = px(0),
      table.border.right.width = px(0)
    ) %>%
    tab_style(
      style = cell_text(align = "center"),
      locations = cells_column_labels(everything())
    )
}

```

### Release Date

<span style="font-size: 90%">**Figure A1**  
*Manual screening outcome by year bin range*</span>

```{r, warning=FALSE, echo=FALSE}
colors <- viridis(10, option="plasma") 
make_removed_counts(lyrics_df, year_bin_range) + 
  scale_fill_manual(
  name = "Screening outcome",  # legend title
  values = c("remove" = colors[5], "keep" = colors[9]),
  labels = c("remove" = "Removed", "keep" = "Kept")) +
  theme(legend.position = "top")

ggsave(here("images", "figure_A1.pdf"))
```
<span style="font-size: 80%"><em>Note.</em> “remove” indicates songs excluded due to manual screening criteria. Bins presented in order of number of songs labelled "Kept". Each row indicates a decade-range year bin, with x axis labels showing the observed metadata range per year bin.</span>

```{r}
table_A1 <- make_apa_freq_table(
  lyrics_df|>rename(`Year Range`= year_bin_range),
  `Year Range`,
  title = md("**Table A1**"),
  subtitle = md("*Frequencies by release year bin*"))

table_A1

gtsave(table_A1, here("tables", "table_A1.tex"))
```
<span style="font-size: 80%"><em>Note.</em> “Removed” indicates songs excluded due to manual screening criteria. Bins presented in reverse order of number of songs removed. Each row indicates a decade-range year bin, with column Year Range showing the observed metadata range per year bin.</span>

### Popularity

Popularity was extremely exponentially distributed, with an extreme left skew. We manually binned to make the quantiles per each of the 7 bins as similar as possible. Thus, the first bin contained the lowest 40% of the population in terms of popularity, while the 7th bin contained the highest 9%:

- **Bin 0:** 1–2 playlist appearances (≤ 40%)
- **Bin 2:** 2–3 appearances (41–53%)
- **Bin 3:** 3–4 appearances (54–60%)
- **Bin 4:** 4–6 appearances (61–70%)
- **Bin 5:** 7–14 appearances (71–80%)
- **Bin 6:** 15–60 appearances (81–91%)
- **Bin 7:** 61–203,345 appearances (top 9% of artists; 91–100%)

<span style="font-size: 90%">**Figure A2**  
*Manual screening outcome by popularity bin percentile range*</span>

```{r}
colors <- viridis(10, option="plasma") 
make_removed_counts(lyrics_df, artist_playlist_frequency_bin_label) + 
  scale_fill_manual(
  name = "Screening outcome",  # legend title
  values = c("remove" = colors[5], "keep" = colors[9]),
  labels = c("remove" = "Removed", "keep" = "Kept")) +
  theme(legend.position = "top")

ggsave(here("images", "figure_A2.pdf"))
```
<span style="font-size: 80%"><em>Note.</em> “remove” indicates songs excluded due to manual screening criteria. Bins presented in reverse order of number of songs labelled "keep".</span>

```{r}
table_A2 <- make_apa_freq_table(
  lyrics_df|>rename(`Artist Popularity`=artist_playlist_frequency_bin_label),
  `Artist Popularity`,
  title = md("**Table A2**"),
  subtitle = md("*Frequencies by artist playlist frequency*"))


table_A2

gtsave(table_A2, here("tables", "table_A2.tex"))
```

<span style="font-size: 80%"><em>Note.</em> “Removed” indicates songs excluded due to manual screening criteria. Bins presented in reverse order of number of songs removed. Each row indicates a popularity percentile range bin, with column Artist Popularity showing the observed percentile range per popularity bin.</span>

### Genre

```{r}
table_A3 <- make_genre_comparison_table(lyrics_df,
  title = md("**Table A3**"),
  subtitle = md("*Comparison of terms of topic tags of initial pool of 15.4k artists, to secondary pool of 2200 songs*")
)  %>%
  cols_width(lda_genre_terms ~ px(350), artist_topic_bin_labels ~ px(300))


table_A3

gtsave(table_A3, here("tables", "table_A3.tex"))
```

@demetriou2024towards  excluded songs if they contained more than >= ~20% non-English words, consisted only of onomatopoeia or nonsense, repeated a single word or phrase, or had too few words. This process yielded 380 songs. We extended screening with the same criteria, resulting in 1,200 songs. We used 800 in the main study and 14 for a pilot. Songs were primarily removed for containing too many non-English words, or having too few or single repeated words. Distribution of included songs by genre bin, as well as results from manual screening procedure are shown in Figure 3. 


<span style="font-size: 90%">**Figure A3**  
*Manual screening outcome by genre*</span>

```{r screening_by_genre}

make_removed_counts(lyrics_df, artist_topic_bin_labels) + 
  scale_fill_manual(
  name = "Screening outcome",  # legend title
  values = c("remove" = colors[5], "keep" = colors[9]),
  labels = c("remove" = "Removed", "keep" = "Kept")) +
  theme(legend.position = "top")

ggsave(here("images", "figure_A3.pdf"))
```

<span style="font-size: 80%"><em>Note.</em> “Remove” indicates songs excluded due to screening criteria.</span>

```{r}
table_A4 <- make_apa_freq_table(
  lyrics_df|>rename(`Artist Topic Tags`=artist_topic_bin_labels),
  `Artist Topic Tags`,
  title = md("**Table A4**"),
  subtitle = md("*Frequencies by artist topic*")
)

table_A4

gtsave(table_A4, here("tables", "table_A4.tex"))
```


#### Topic

<span style="font-size: 90%">**Figure A4**  
*Manual screening outcome by lyric topic*</span>

```{r}
make_removed_counts(lyrics_df, topic_label) + 
  scale_fill_manual(
  name = "Screening outcome",  # legend title
  values = c("remove" = colors[5], "keep" = colors[9]),
  labels = c("remove" = "Removed", "keep" = "Kept")) +
  theme(legend.position = "top")

ggsave(here("images", "figure_A4.pdf"))
```

## Political Speech Stimuli


Repository of sampling procedure can be found here: https://github.com/Gargant0373/PresidentialSpeechSampling/blob/master/sample.py


<span style="font-size: 90%">**Figure A5**  
*Number of speech excerpts per President*</span>

```{r}
colors <- viridis(10) 

speech_df |> 
  
  filter(id %in% speech_IDs) |>
  
  count(president, party, name = "n") |>
  
  mutate(president = fct_reorder(president, n, .desc = F)) |>
  
  ggplot(aes(x = president, y = n, fill = party)) +
    geom_col(alpha = .75) +
    coord_flip() +
    labs(x = NULL, y = "Count") +
    theme_apa() + 
  scale_fill_manual(
  name = "Party",  
  values = c("Democratic" = colors[5], "Republican" = colors[9]),
  labels = c("remove" = "Removed", "keep" = "Kept")) +
  theme(legend.position = "top")

ggsave(here("images", "figure_A5.pdf"))
```


```{r}
df <- speech_df

# function to summarize by party
summarise_party <- function(data) {
  data %>%
    group_by(party) %>%
    summarise(
      n_excerpts    = n(),
      n_speeches    = n_distinct(url),
      n_presidents  = n_distinct(president),
      .groups = "drop"
    ) %>%
    mutate(
      excerpts_per_president = n_excerpts / n_presidents,
      pct_excerpts = n_excerpts / sum(n_excerpts))}

# 1) Full pool
full_summary <- summarise_party(df) %>%
  mutate(sample = "Initial pool")

# 2) Annotated subsample
sub_df <- df %>% filter(id %in% speech_IDs)
sub_summary <- summarise_party(sub_df) %>%
  mutate(sample = "Sample")

# 3) Combine
combined_summary <- bind_rows(full_summary, sub_summary) %>%
  relocate(sample, .before = party)

# 4) Render as gt table
table_A5 <- combined_summary %>%
  gt(groupname_col = "sample") %>%   
    tab_header(
    title = md("**Table A5**"),   
    subtitle = md("*Descriptive statistics of speech excerpts by party*")) |>
  cols_label(
    party = "Party",
    n_excerpts = "Total excerpts",
    n_speeches = "Unique speeches",
    n_presidents = "Unique presidents",
    excerpts_per_president = "Avg. excerpts per president",
    pct_excerpts = "% of all excerpts") %>%
  
  fmt_number(columns = c(n_excerpts, n_speeches, n_presidents), decimals = 0) %>%
  fmt_number(columns = excerpts_per_president, decimals = 1) %>%
  fmt_percent(columns = pct_excerpts, decimals = 1) %>%
  
  tab_options(
    table.font.size = 12,
    heading.align = "left",    
    table_body.hlines.width = px(0),
    table.border.top.width = px(0),
    table.border.bottom.width = px(1),
    column_labels.border.top.width = px(0),
    column_labels.border.bottom.width = px(1)) %>%
  
  tab_style(
    style = cell_text(align = "center", whitespace = "pre-wrap"),
    locations = cells_column_labels(everything())) |>
  
  cols_width(
    party ~ px(100),
    n_excerpts ~ px(100),
    n_speeches ~ px(100),
    n_presidents ~ (100), 
    excerpts_per_president ~ px(100),
    pct_excerpts ~ px(100)) %>%
    
  tab_style(
    style = cell_text(align = "center"),
    locations = cells_row_groups()) %>%
  
  cols_align(align = "center", columns = everything())


table_A5

gtsave(table_A5, here("tables", "table_A5.tex"))
``` 

## Word Counts

```{r}
lyrics_summary <- lyrics_df |> filter(item_ID %in% lyrics_IDs) |> 
  summarize(mean = mean(word_count), sd = sd(word_count))

speech_summary <- speech_df |> filter(id %in% speech_IDs) |> 
  summarize(mean = mean(word_count), sd = sd(word_count))
```

<span style="font-size: 90%">**Figure 6**  
*Density plot of word counts in speech and lyric excerpts*</span>
```{r}
df <- bind_rows(
  lyrics_df |> filter(item_ID %in% lyrics_IDs) |> 
    select(item_ID, word_count) |> mutate(source = "Lyrics", item_ID = as.character(item_ID)), 
  speech_df |> filter(id %in% speech_IDs) |> 
    select(id, word_count) |> rename(item_ID = id) |>
    mutate(source = "Speeches")) |> rename(Source = source)

df |>
  
  ggplot(aes(x = word_count, fill = Source)) +
    geom_density(alpha = 0.7) +
  scale_fill_manual(values = c("Speeches" = "#21908C", "Lyrics" = "#440154")) +
    labs(
    x = "Word count",
    y = "Density") +
  theme_apa()

ggsave(here("images", "figure_A6.pdf"))
```

